\documentclass[twoside,twocolumn,10pt]{article}

% -----------------------------
% Packages
% -----------------------------
\usepackage{fancyhdr}
\usepackage{geometry}
\usepackage{abstract}
\usepackage{graphicx}
\usepackage{titlesec}
\usepackage{ragged2e}
\usepackage{amssymb}
\usepackage{parskip}
\usepackage{amsmath}
\usepackage{newtxtext}
\usepackage{booktabs}
\usepackage{array}
\usepackage{balance}
\usepackage[backend=bibtex,style=ieee]{biblatex}
\addbibresource{references.bib}
\usepackage[font=footnotesize,labelfont=bf,justification=raggedright,format=plain]{caption}
\usepackage{float}

% -----------------------------
% Bibliography Heading
% -----------------------------
\defbibheading{bibliography}[\refname]{%
  \section{\MakeUppercase{REFERENCES}}}

% -----------------------------
% Custom Column Types
% -----------------------------
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}

% -----------------------------
% Caption Formatting
% -----------------------------
\captionsetup[table]{labelsep=period, justification=raggedright}
\captionsetup[figure]{labelsep=period, justification=raggedright}

% -----------------------------
% Page Geometry
% -----------------------------
\geometry{
    a4paper,
    left=0.8in,
    right=0.8in,
    top=1in,
    bottom=1in
}

\setlength{\columnsep}{15pt}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0pt}
\hyphenpenalty=10000
\exhyphenpenalty=10000
\frenchspacing

% -----------------------------
% Abstract Formatting
% -----------------------------
\renewcommand{\abstractname}{\hspace{1.3em}\textbf{Abstract}}
\renewcommand{\abstractnamefont}{\normalfont\fontsize{10}{14}\bfseries\MakeUppercase}
\renewcommand{\absnamepos}{flushleft}
\renewcommand{\abstracttextfont}{\normalfont\fontsize{11}{13}\selectfont}

% -----------------------------
% Header / Footer Style
% -----------------------------
\setcounter{page}{1}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[C]{\small RESOURCE-CONSTRAINED MULTIMODAL DISASTER ASSESSMENT}
\fancyhead[CO]{\small Zohaifa et al. (2026)}
\fancyfoot[C]{\thepage}

% -----------------------------
% First Page Style
% -----------------------------
\fancypagestyle{firstpage}{
  \fancyhf{}
  \fancyhead[LO]{\small Volume X, Issue X, Year}
 \fancyhead[RO]{\small LGU J. Comput. Sci. \& IT}
  \fancyfoot[L]{\footnotesize Received: DD Month YYYY; Revised: DD Month YYYY; Accepted: DD Month YYYY\\\textsuperscript{*}These authors contributed equally to this work.\\Corresponding Authors: A. Zohaifa, J. Uddin, M. Rahman}
  \fancyfoot[R]{\footnotesize \textcopyright\ Year by the Author(s). Published by LGURJCSIT,\\ \hspace{0.5cm} under the CC-BY license}
  \renewcommand{\headrulewidth}{0.4pt}
  \renewcommand{\footrulewidth}{0.4pt}
}

% -----------------------------
% Title / Author
% -----------------------------
\title{\fontsize{16}{20}\selectfont\textbf{RESOURCE-CONSTRAINED MULTIMODAL DISASTER ASSESSMENT: A NOISE-RESILIENT FUSION ARCHITECTURE FOR LOW-RESOURCE AND HIGH-RESOURCE ENVIRONMENTS}}

\author{
    \fontsize{12}{14}\selectfont
    Abu Zohaifa\textsuperscript{1*}, Jalal Uddin\textsuperscript{1*}, Mushfiqur Rahman\textsuperscript{1*}\\[1ex]
    \fontsize{12}{14}\selectfont
    \textsuperscript{1}Department of Computer Science and Engineering, International Islamic University Chittagong (IIUC), Bangladesh
}

\date{}

% -----------------------------
% Section Formatting
% -----------------------------
\titleformat{\section}{\normalsize\bfseries\uppercase}{\thesection.}{1em}{}
\titleformat{\subsection}{\normalsize\bfseries\flushleft}{\thesubsection.}{1em}{}
\titleformat{\subsubsection}{\normalsize\bfseries\itshape\flushleft}{\thesubsubsection.}{1em}{}

% -----------------------------
% Document Start
% -----------------------------
\begin{document}

\twocolumn[
\begin{@twocolumnfalse}

\vspace{-0.7cm}
\noindent
\raggedright
{\fontsize{11}{13}\selectfont Original Article}
\hfill
{\fontsize{11}{13}\selectfont\footnotesize https://doi.org/xx.xxxx/xxxx}

\vspace{-0.5cm}

\maketitle
\thispagestyle{firstpage}

\section*{\hspace{-1.5mm}Abstract}
\fontsize{10}{12}\selectfont
\justifying
\noindent
Multimodal disaster assessment using Unmanned Aerial Vehicle (UAV) imagery and social media text is a critical technological frontier for rapid emergency response, resource allocation, and victim localization. However, deploying such deep learning models in resource-constrained environments—particularly involving linguistically low-resource languages like Bengali—poses significant operational challenges. These include handling massive unstructured noise in crowd-sourced text, severe long-tail class imbalance, and the computational limitations of edge devices. To address these compounding issues, we propose a highly optimized, noise-resilient dual-encoder late-fusion architecture. The vision pathway utilizes a lightweight Swin-Tiny backbone enhanced with aggressive photometric augmentations, while the semantic pathway leverages state-of-the-art text encoders (BanglaBERT for local contexts, and XLM-RoBERTa for global generalization). Crucially, we introduce a Targeted Fast Gradient Method (FGM) adversarial training protocol restricted exclusively to word embeddings, which heavily regularizes the text manifold without corrupting visual-spatial representations. To rectify log-bias shifts induced by inherent class imbalances, our framework integrates a Dynamic Pseudo-Labeling Strategy (DIPS) and an Optuna-optimized Level-2 LightGBM stacker. Extensive benchmarking on the regional BanglaCalamityMMD and the global CrisisMMD v2.0 datasets reveals that our robust pipeline achieves a Macro F1 score of 0.745 in resource-scarce conditions, significantly outperforming conventional unimodal and baseline multimodal architectures.

\vspace{0.3cm}
\noindent\textbf{Keywords:}
Multimodal Disaster Assessment, Dual-Encoder Architecture, Fast Gradient Method, Pseudo-Labeling, Low-Resource Languages, UAV Surveillance

\vspace{0.8cm}
\end{@twocolumnfalse}
]

% -----------------------------
% Main Text
% -----------------------------
\section{Introduction}
\fontsize{11}{13}\selectfont

In the immediate aftermath of a catastrophic natural or human-induced disaster, the velocity and accuracy of severity assessment directly dictate the success of emergency response protocols. Real-time situational awareness is increasingly derived from two primary, yet highly divergent, data streams: Unmanned Aerial Vehicle (UAV) visual surveillance and crowd-sourced social media text. Combining these modalities provides a holistic view of the event, fusing the macro-level geographic destruction captured by drones with the micro-level human context broadcasted by victims and witnesses.

Despite its theoretical advantages, multimodal disaster assessment remains a profoundly complex engineering problem. Social media streams are notoriously noisy, featuring typographical errors, slang, and unverified claims. Furthermore, in developing nations where disasters strike with devastating frequency (e.g., severe flooding and cyclones in Bangladesh), the local languages are often computationally ``low-resource." Standard transformer-based architectures frequently suffer from routing collapse and catastrophic overfitting when processing low-resource languages like Bengali due to inadequate pre-training corpora. Similarly, UAV imagery often suffers from extreme photometric variances—such as sun glare, overcast shadows, and sensor noise—making visual feature extraction highly unstable. 

In this work, we present a resource-constrained, robust dual-encoder multimodal fusion architecture specifically engineered to mitigate these compounding sources of noise and data scarcity. Our contributions are distinctly four-fold:
\begin{itemize}
    \item We design a highly efficient late-fusion architecture bridging a computationally lightweight Swin-Tiny vision transformer with domain-specific BERT language models.
    \item We formulate and deploy a \textit{Targeted Fast Gradient Method (FGM)}, restricting adversarial perturbations exclusively to the text embedding space to prevent visual manifold degradation.
    \item We address extreme class imbalances using a Dynamic Pseudo-Labeling Strategy (DIPS) coupled with Level-2 LightGBM stacking optimized by Optuna for log-bias correction.
    \item We validate the framework's scalability, proving its high efficacy on the Bengali-centric BanglaCalamityMMD dataset (Phase 1) and demonstrating seamless global generalization on the multilingual CrisisMMD v2.0 dataset (Phase 2).
\end{itemize}

\section{Methodology}
\fontsize{11}{13}\selectfont

To effectively parse the disjointed semantic and visual disaster manifolds, our architecture maintains independent feature pathways before executing a dense late-stage projection.

\subsection{Multimodal Dual-Encoder Late Fusion}
The visual representation pathway extracts spatial hierarchies using the \texttt{swin\_tiny\_patch4\_window7\_224} architecture. Swin Transformers compute self-attention within shifted local windows, generating a hierarchical feature map at a fraction of the computational complexity of standard Vision Transformers. To combat the scarcity of disaster variations, we apply aggressive Hue, Saturation, and Value (HSV) geometric distortions during training. 

The semantic pathway is powered by \texttt{csebuetnlp/banglabert} for localized Bengali tasks, and \texttt{xlm-roberta-large} for global generalization. Both encoders process the text into a comprehensive pooled representation. The output arrays are concatenated to form the dense multi-modal feature vector $f_{mm}$:
\begin{equation}
    f_{mm} = \text{Concat}(f_{vision}, f_{text})
\end{equation}
This feature vector is passed through a regularizing dropout layer ($p=0.3$) and mapped to the $k$ severity classes using a linear classification head: $z = W_f f_{mm} + b_f$.

\subsection{Targeted Adversarial Training (FGM)}
Standard adversarial training techniques like the Fast Gradient Method (FGM) globally perturb all trainable parameters. In a highly-augmented multimodal setting, globally applying FGM often destroys the visual spatial coherence established by the Swin-Tiny backbone. Therefore, we introduce a targeted adversarial block. We strictly calculate the adversarial noise $r_{adv}$ over the text encoder's \texttt{word\_embeddings}:
\begin{equation}
    r_{adv} = \epsilon \frac{\nabla_e \mathcal{L}(e, y)}{||\nabla_e \mathcal{L}(e, y)||_2}
\end{equation}
where $e$ represents the word embeddings and $\epsilon$ is the scaling factor. This creates a powerful regularizing effect on the textual manifold, forcing the model to become mathematically resilient to social media typos and dialectical noise without destabilizing the visual weights.

\subsection{FP16 Scaling and Gradient Anomaly Check}
Due to the stringent memory constraints of edge-deployable hardware, we utilize FP16 Mixed Precision training. Multimodal optimization often experiences transient exploding gradients, causing the dynamic loss scaler to collapse to \texttt{NaN}. To ensure training continuity, we deploy a gradient anomaly safeguard. If the scaler shrinks consecutively ($\text{scale}_{t} < \text{scale}_{t-1}$), the optimizer and scheduler steps are explicitly skipped, allowing the gradients to stabilize dynamically.

\subsection{Dynamic Pseudo-Labeling Strategy (DIPS)}
Real-world disaster datasets possess enormous long-tail class imbalances. To augment our training distribution organically, we implement DIPS over unannotated test streams. DIPS monitors the softmax prediction distributions across three sequential epochs. An unannotated sample $x_i$ is promoted to the pseudo-ground-truth set $D_{pseudo}$ only if its mean confidence exceeds a strict threshold ($\mu_{conf} \ge 0.85$) while maintaining minimal epistemic uncertainty ($\sigma_{conf} \le 0.15$). This vetted data acts as an anchor during the final fine-tuning phase.

\subsection{Optuna-Calibrated LightGBM Stacking}
Finally, the standard Cross-Entropy loss naturally biases predictions toward the majority class. We resolve this log-bias shift post-training using a stacking framework. We extract Out-Of-Fold (OOF) multimodal features from the validation partitions and train a Level-2 LightGBM classifier. We utilize Optuna to perform an extensive Bayesian hyperparameter search over the LightGBM learning rate, leaf volume, and boosting rounds, optimizing directly for the Macro F1-score.

\section{Experiments and Results}
\fontsize{11}{13}\selectfont

The architecture was benchmarked in two distinct phases targeting regional low-resource settings and global generalization.

\subsection{Phase 1: BanglaCalamityMMD (Low-Resource)}
We conducted a rigorous ablation study on the local BanglaCalamityMMD dataset to isolate the performance gains yielded by each architectural component. As presented in Table \ref{table1}, the base multimodal fusion architecture (Run A) struggles severely with noise, yielding a Macro F1 of 0.692. The integration of Targeted FGM (Run B) dramatically smooths the embedding space, jumping the F1-score to 0.724. Finally, incorporating DIPS pseudo-labeling and the Level-2 Optuna stacker (Run C) successfully realigns the class biases, establishing a state-of-the-art threshold of 0.745.

% -----------------------------
% Table 1
% -----------------------------
\begin{table}[h]
\centering
\caption{Ablation Study on BanglaCalamityMMD (Phase 1)}
\label{table1}
\renewcommand{\arraystretch}{1.2}
\setlength{\tabcolsep}{6pt}
\resizebox{\columnwidth}{!}{
\begin{tabular}{p{4cm} p{2cm} p{2cm}}
\toprule
\textbf{Architectural Configuration} & \textbf{Macro F1} & \textbf{Accuracy} \\
\midrule
Run A: Base Fusion Model & 0.692 & 0.761 \\
Run B: Base + Targeted FGM & 0.724 & 0.793 \\
Run C: Full (FGM + DIPS + Stacking) & \textbf{0.745} & \textbf{0.812} \\
\bottomrule
\end{tabular}
}
\end{table}

\subsection{Phase 2: CrisisMMD v2.0 (High-Resource)}
A true disaster-response framework must perform reliably across geographic and linguistic borders. We subjected the exact architecture (substituting the text backbone for the multilingual XLM-RoBERTa-Large) to the CrisisMMD v2.0 benchmark targeting the complex `Damage Severity' task. 

% -----------------------------
% Table 2
% -----------------------------
\begin{table}[h]
\centering
\caption{Generalization Benchmark on CrisisMMD (Phase 2)}
\label{table2}
\renewcommand{\arraystretch}{1.2}
\setlength{\tabcolsep}{6pt}
\resizebox{\columnwidth}{!}{
\begin{tabular}{p{4cm} p{2cm} p{2cm}}
\toprule
\textbf{Dataset Partition} & \textbf{Macro F1} & \textbf{Accuracy} \\
\midrule
Hurricane Harvey (Texas) & 0.751 & 0.820 \\
Hurricane Irma (Florida) & 0.748 & 0.817 \\
\midrule
\textbf{Global Pipeline Average} & \textbf{0.749} & \textbf{0.818} \\
\bottomrule
\end{tabular}
}
\end{table}

The empirical results in Table \ref{table2} confirm exceptional generalization capability. Our model securely maintains a $\sim0.75$ Macro F1 score across heavily distinct geographic catastrophes without any architectural restructuring, verifying the universal robustness of the Targeted FGM and DIPS algorithms.

\section{Conclusion}
\fontsize{11}{13}\selectfont

This paper presented an end-to-end, resource-constrained multimodal architecture designed explicitly for extreme disaster severity assessment. By intelligently pairing a lightweight shifted-window vision transformer (Swin-Tiny) with robust local and global language encoders, we forged a highly effective late-fusion pipeline. The introduction of Targeted FGM exclusively protected the semantic manifold from social media noise, while DIPS and LightGBM stacking mitigated severe long-tail class imbalances. Comprehensive experiments unequivocally demonstrated that our framework systematically elevates performance in heavily constrained, low-resource Bengali environments while scaling flawlessly to high-resource international benchmarks. The resulting pipeline offers a crucial, reproducible blueprint for deploying resilient AI models directly onto edge and UAV hardware during critical planetary emergencies.

\vspace{0.5cm}

\small
\textbf{Conflict of Interest:} The authors declare no conflict of interest.

\textbf{Author Contributions:} These authors contributed equally to this work.

\textbf{Funding:} This research received no external funding.

\textbf{Ethical Statement:} This study follows ethical and academic integrity guidelines.

\balance
\fontsize{11}{13}\selectfont
\printbibliography

\end{document}
